Hot and fresh: Pervasive climate stressors of seagrass in a large Gulf coast estuary

### 2023-11-06 manu-draft.qmd
update all figs, tables, results with 2022 fim data

Increasing water temperature and decreasing salinity generally followed the meteorological trends for all three *in situ* datasets (EPC, FIM, and PDEM, @fig-meteowqraw, @tbl-temptrndtab, @tbl-saltrndtab, figures in supplement). Note that for tables [-@tbl-temptrndtab] and [-@tbl-saltrndtab], comparable time periods were evaluated between the datasets when possible given the different sample sizes, and therefore power, to detect trends. The strongest trends were observed for the EPC dataset which had the longest record from 1975 to 2022. The top and bottom water temperature or salinity changes were similar across bay segments likely because of the relatively shallow water depths and minimal stratification in Tampa Bay (@fig-meteowqraw). Trends in water temperature were similar across bay segments with significant increases varying from 0.3 to 0.4 $^\circ$C per year, with a total increase of 1.4 (OTB) to 2.0 (HB) $^\circ$C from 1975 to present (@tbl-temptrndtab). Salinity trends were similar between bay segments, although overall salinity was of course higher for bay segments closer to the Gulf of Mexico. Only Old Tampa Bay and Lower Tampa Bay had significantly decreasing trends, with decreases of 0.06 (OTB) and 0.04 (LTB) ppt per year, with a total decrease of 2.7 (OTB) and 2.0 (LTB) ppt from 1975 to present. As expected, intra-annual variability of water temperature was much higher than salinity following the larger seasonal variation in temperature (@fig-meteowqraw). This variation may also be related to differences among the monitoring stations in each segment because the averages combine monthly and spatial variation. Significant increasing temperature trends were also observed for the FIM dataset from 1996 to present in OTB and from 2004 to present in HB (@tbl-temptrndtab). No significant temperature or salinity trends were observed in the other bay segments for the FIM and PDEM datasets from 1996 to present nor 2004 to present (see figures in supplement). However, the EPC dataset showed significant increasing temperature across all time periods and bay segments. Salinity trends from the EPC dataset were not significant from 1996 to present, but were significantly decreasing from 2004 to present in all bay segments (@tbl-saltrndtab).

>>>>

Increasing water temperature and decreasing salinity generally followed the meteorological trends for all three *in situ* datasets (EPC, FIM, and PDEM, @fig-meteowqraw, @tbl-temptrndtab, @tbl-saltrndtab, figures in supplement). Note that for tables [-@tbl-temptrndtab] and [-@tbl-saltrndtab], comparable time periods were evaluated between the datasets when possible given the different sample sizes, and therefore power, to detect trends. The strongest trends were observed for the EPC dataset which had the longest record from 1975 to 2022. The top and bottom water temperature or salinity changes were similar across bay segments likely because of the relatively shallow water depths and minimal stratification in Tampa Bay (@fig-meteowqraw). Trends in water temperature were similar across bay segments with significant increases varying from 0.3 to 0.4 $^\circ$C per year, with a total increase of 1.4 (OTB) to 2.0 (HB) $^\circ$C from 1975 to present (@tbl-temptrndtab). Salinity trends were similar between bay segments, although overall salinity was of course higher for bay segments closer to the Gulf of Mexico. Only Old Tampa Bay and Lower Tampa Bay had significantly decreasing trends, with decreases of 0.06 (OTB) and 0.04 (LTB) ppt per year, with a total decrease of 2.7 (OTB) and 2.0 (LTB) ppt from 1975 to present. As expected, intra-annual variability of water temperature was much higher than salinity following the larger seasonal variation in temperature (@fig-meteowqraw). This variation may also be related to differences among the monitoring stations in each segment because the averages combine monthly and spatial variation. Significant increasing temperature trends were also observed for the FIM dataset from 1996 to present in OTB and from 2004 to present in OTB and HB (@tbl-temptrndtab). No significant temperature or salinity trends were observed in the other bay segments for the FIM and PDEM datasets from 1996 to present nor from 2004 to present (see figures in supplement). However, the EPC dataset showed significant increasing temperature across all time periods and bay segments. Salinity trends from the EPC dataset were not significant from 1996 to present, but were significantly decreasing from 2004 to present in all bay segments (@tbl-saltrndtab).
